---
title: "Strings and Factors"
author: "Vanessa Dinh"
output: github_document
---

```{r}
library(rvest)
library(p8105.datasets)
library(tidyverse)
library(stringr)
library(forcats)
```
```{r}
string_vec = c("my", "name", "is", "jeff")

str_detect(string_vec, "jeff")
```
```{r}
str_replace(string_vec, "jeff", "Jeff")
```

Designate matches at beginning with carrot
```{r}
string_vec = c(
  "i think we all rule for participating",
  "i think i have been caught",
  "i think this will be quite fun actually",
  "it will be fun, i think"
  )

str_detect(string_vec, "^i think")
```
Matches at end with dollar sign

```{r}
str_detect(string_vec, "i think$")
```
Designate list of characters with brackets
```{r}
string_vec = c(
  "Y'all remember Pres. HW Bush?",
  "I saw a green bush",
  "BBQ and Bushwalking at Molonglo Gorge",
  "BUSH -- LIVE IN CONCERT!!"
  )

str_detect(string_vec,"[Bb]ush")
```

Range of characters with brackets and dash in between
```{r}
string_vec = c(
  '7th inning stretch',
  '1st half soon to begin. Texas won the toss.',
  'she is 5 feet 4 inches tall',
  '3AM - cant sleep :('
  )

str_detect(string_vec, "^[0-9][a-zA-Z]")
```

Character . matches anything
```{r}
string_vec = c(
  'Its 7:11 in the evening',
  'want to go to 7-11?',
  'my flight is AA711',
  'NetBios: scanning ip 203.167.114.66'
  )

str_detect(string_vec, "7.11")
```

Some characters are “special”. These include [ and ], ( and ), and .. If you want to search for these, you have to indicate they’re special using \. Unfortunately, \ is also special, so things get weird.
```{r}
string_vec = c(
  'The CI is [2, 5]',
  ':-]',
  ':-[',
  'I found the answer on pages [6-7]'
  )

str_detect(string_vec, "\\[")
```

## Why factors are weird

```{r}
factor_vec = factor(c("male", "male", "female", "female"))

as.numeric(factor_vec)

factor_vec = fct_relevel(factor_vec, "male")
as.numeric(factor_vec)
```
## NSDUH (National Survey on Drug Use and Health)

```{r}
nsduh_url = "http://samhda.s3-us-gov-west-1.amazonaws.com/s3fs-public/field-uploads/2k15StateFiles/NSDUHsaeShortTermCHG2015.htm"

table_marj = 
  read_html(nsduh_url) %>% 
  html_table() %>% 
  first() %>%
  slice(-1)

#slice gets rid of first row in data
```
tidy up NSDUH data, one backslash is a special character, need two to designate other special characters, exclamation point for NOT
```{r}
marj_df = 
  table_marj %>% 
  select(-contains("P value")) %>% 
  pivot_longer(
    -State,
    names_to = "age_year",
    values_to = "percent"
  ) %>% 
  mutate(
    percent = str_replace(percent, "[a-b]$", ""),
    percent = as.numeric(percent)
  ) %>% 
  separate(age_year, into = c("age", "year"), sep = "\\(") %>% 
  mutate(
    year = str_replace(year, "\\)", "")
  ) %>% 
  filter(
    !(State %in% c("Total U.S.", "Northeast", "Midwest", "South", "West", "District of Columbia")))

```


```{r}
marj_df %>% 
  filter(age == "12-17") %>% 
  mutate(State = fct_reorder(State, percent)) %>% 
  ggplot(aes(x = State, y = percent, color = year)) + 
    geom_point() + 
    theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

## Restaurant Inspections
```{r}
data("rest_inspec")
```
```{r}
rest_inspec %>% 
  slice(1:1000) %>% view()
```
```{r}
rest_inspec %>% 
  group_by(boro, grade) %>% 
  summarize(n_obs=n()) %>% 
  pivot_wider(
    names_from = grade,
    values_from = n_obs
  )
```

```{r}
rest_inspec = 
  rest_inspec %>% 
  filter(grade %in% c("A", "B", "C"), boro != "Missing") %>% 
  mutate(boro = str_to_title(boro))

rest_inspec %>% 
  group_by(boro, grade) %>% 
  summarize(n_obs=n()) %>% 
  pivot_wider(
    names_from = grade,
    values_from = n_obs
  )
```

Find pizza places

```{r}
rest_inspec %>% 
  filter(str_detect(dba, "Pizza")) %>% 
  group_by(boro, grade) %>% 
  summarize(n = n()) %>% 
  pivot_wider(names_from = grade, values_from = n)
```

```{r}
rest_inspec %>% 
  filter(str_detect(dba, "[Pp][Ii][Zz][Zz][Aa]")) %>% 
  group_by(boro, grade) %>% 
  summarize(n = n()) %>% 
  pivot_wider(names_from = grade, values_from = n)
```

```{r}
rest_inspec %>% 
  filter(str_detect(dba, "[Pp][Ii][Zz][Zz][Aa]")) %>%
  ggplot(aes(x = boro, fill = grade)) + 
  geom_bar() 
```

```{r}
rest_inspec %>% 
  filter(str_detect(dba, "[Pp][Ii][Zz][Zz][Aa]")) %>%
  mutate(boro = fct_infreq(boro)) %>%
  ggplot(aes(x = boro, fill = grade)) + 
  geom_bar()
```
reconverted to string variable with str_replace after fct_infreq

```{r}
rest_inspec %>% 
  filter(str_detect(dba, "[Pp][Ii][Zz][Zz][Aa]")) %>%
  mutate(
    boro = fct_infreq(boro),
    boro = str_replace(boro, "Manhattan", "The City")) %>%
  ggplot(aes(x = boro, fill = grade)) + 
  geom_bar() 
```
using replace() tries to create a new factor variable, so use fct_recode()
```{r}
rest_inspec %>% 
  filter(str_detect(dba, "[Pp][Ii][Zz][Zz][Aa]")) %>%
  mutate(
    boro = fct_infreq(boro),
    boro = replace(boro, which(boro == "Manhattan"), "The City")) %>%
  ggplot(aes(x = boro, fill = grade)) + 
  geom_bar()
```
```{r}
rest_inspec %>% 
  filter(str_detect(dba, regex("pizza", ignore_case = TRUE))) %>%
  mutate(
    boro = fct_infreq(boro),
    boro = fct_recode(boro, "The City" = "Manhattan")) %>%
  ggplot(aes(x = boro, fill = grade)) + 
  geom_bar()
```


